---
title: "Practical on mechanistic recruitment models"
author: "Bitetto Isabella - COISPA"
date: "2024-11-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

To test the significance of environmental variables on recruitment process, first we need to select meaningful covariates. 
The environmental time series used in this tutorial have been derived from Copernicus (https://www.copernicus.eu/en, reanalysis approach on Bernhard presentation)

The variables here considered are: 

+ the sea surface temperature (SST), 
+ bottom temperature (botT), 
+ salinity (so), 
+ bottom salinity (botso) 
+ net primary production value (nppv). 

The environmental data are considered on a depth range based on the distribution of each species in the study area: 

+ European hake: 50-500 m; 
+ red mullet: 0-200 m; 
+ deep-water rose shrimp:50-500 m; 
+ red shrimps: 200-800 m 

and grouped in GSAs based on the stock unit considered. Monthly values of the environmental variables were averaged yearly for each depth range considered. 

The environmental data related to the past and present years (hind cast) span from 1987 to 2021 while the projections cover the years 2006-2099. 

Also for the projections, the environmental data were obtained from 3D monthly POLCOMS-ERSEM dataset with a resolution of 0.1° (approximately 10 km) for two RCP (Representative Concentration Pathways) climate scenarios: 
+ a moderate one, the RCP4.5 
+ the RCP8.5 (no climate protection) as the highest baseline emissions scenario in which emissions continue to rise throughout the twenty-first century (i.e. until 2099).


## Required packages to run this tutorial

In order to execute the code in this tutorial you should have the following packages installed:

CRAN: modelr, nlstools, MuMIn, rlist, formula.tools, corrplot


```{r packages, echo=TRUE, warning = FALSE, message = FALSE,results=FALSE}

#install.packages(c("modelr","nlstools","MuMIn","rlist","formula.tools","corrplot"),repos="https://cloud.r-project.org")

library(modelr)
library(nlstools)
library(MuMIn)
library(rlist)
library(formula.tools)
library(corrplot)

```

## Loading the datasets

The second step is the load the data for fitting the recruitment models:
+ pairs recruits-SSB from stock assessment models. In this example we will use the last stock assessment results of European hake in Adriatic Sea, publicly available https://www.fao.org/gfcm/data/star/en/;
+ environmental variables: the time series of sst (sea surface temperature), botT (bottom temperature), sal (salinity), botsal (bottom salinity) and nppv (net primary production) was derived from Copernicus spatial data, averaging the grid cells values between 50 and 500 m (preferred habitat of the species). 
Both data sets are also available here: https://figshare.com/s/6ffe7e46031b4990894e

```{r data, echo=TRUE, warning = FALSE, message = FALSE}
load("Recruitment_tutorial_data.Rdata")

head(data)

load("env_hindcast.Rdata")

head(env_hindcast)

case="HKE19"
```

# Exploring correlations among covariates

Correlation plots provide a quick, intuitive way to assess relationships at a glance.Our objective in this case is to avoid including correlated variables in the model fitting process to minimize redundancy and reduce the risk of spurious relationships. 
For European hake in the Western Ionian SEa, we observe that sea surface temperature (SST) is highly correlated with bottom temperature, and bottom salinity is correlated with all variables except nppv. Therefore, we will explore models that include one of the correlated variables at a time to ensure a more robust analysis.

```{r correlation, echo=TRUE, warning = FALSE, message = FALSE}

data= data[data$Year %in% unique(env_hindcast$YEAR),] 
y=data[,c(1)]
R=as.numeric(data[,c(2)])
S=as.numeric(data[,c(3)])

 

botT<-env_hindcast$botT
sst <-env_hindcast$sst
so  <-env_hindcast$so
botso<-env_hindcast$botso
nppv<-env_hindcast$nppv

varr=cbind(botT,sst,so,botso,nppv)[-c(1:5),]
M = cor(varr)
corrplot(M, method = 'number',type="full",number.cex=1.7,tl.cex=1.7,cl.cex=1.5,na.rm=TRUE)

```

# Fitting Environmentally Mediated Stock-Recruitment models

## Beverton and Holt

In this part of the code the years where the environmental information is available are selected. Then,  we fit the models starting from seed values a and b of the Beverton and Holt SRR.
Initial values can be derived from previous estimates, stock assessment or by literature.

We first fit the classical model without environmental variabes that we will define as naive model.

```{r BH classical model, echo=TRUE, warning = FALSE, message = FALSE}

ain=86.69208	
bin=5.99E-04

data_case<-data.frame(cbind(data,botT,sst,so,botso,nppv))
colnames(data_case)[1:3]<-c("year","R","S")
#data_case<-cbind(data_case,env_hindcast$botT,env_hindcast$sst,env_hindcast$so,env_hindcast$botso,env_hindcast$nppv)

N=length(y)
table_res<-c()

R=as.numeric(data_case$R)
S=as.numeric(data_case$S)
y=as.numeric(data_case$year)
  


mod<-nls(R~ a*S/(1+b*S), start=list(a=ain,b=bin) )
err<-residuals(mod)
N_nona<-N-sum(length(which(is.na(err))))
SSE=sum(err^2,na.rm=T)
ME=sum(abs(err),na.rm=T)/N_nona
MSE=SSE/N_nona+ME^2

table_res<-rbind(table_res, c(
  paste("BH"),
  as.character(formula(mod)) ,
  coeffs(mod),"","", "",AIC(mod) , SSE, ME, MSE) )

modR<-list(classic=mod)


```

The second step is to fit the models with one covariate at a time. We will take advantage of the coefficients of the naive model and we will use them as initial values for the next models.
With salinity and bottom salinity the model is not converging.

```{r BH model with one covariate, echo=TRUE, warning = FALSE, message = FALSE}

ain<-coeffs(mod)[1]
bin<-coeffs(mod)[2]
cin<-0.1
din<-0.1
fin<-0.1

data_case$S<-as.numeric(data_case$S)
data_case$R<-as.numeric(data_case$R)

plot(y,R,lwd=1,pch=19,xlab="Year",ylab="Recruits",ylim=c(min(R)*0.9,max(R)*2))
lines(y,predict(modR[[1]]),lwd=3,col="black")
title(paste(case, "-hindcast"))
legend_names<-"BH"
legend_color<-"black"
legend_lty<-1

#botT
mod<-nls(R~a*S/(1+b*S)*exp(-c*botT), start=list(a=ain,b=bin,c=cin),data=data_case )
err<-residuals(mod)
N_nona<-N-sum(length(which(is.na(err))))
SSE=sum(err^2,na.rm=T)
ME=sum(abs(err),na.rm=T)/N_nona
MSE=SSE/N_nona+ME^2
table_res<-rbind(table_res, c(
  paste("BH botT"),
  as.character(formula(mod)),coeffs(mod),"","", AIC(mod) , SSE, ME, MSE) )
modR<-list.append(modR, botT=mod)
lines(y,predict(mod),lwd=3,col=3)
legend_color<-c(legend_color,3)
legend_names<-c(legend_names, "BH botT")
legend_lty<-c(legend_lty,1)


#sst
mod<-nls(R~ a*S/(1+b*S)*exp(-c*sst), start=list(a=ain,b=bin,c=cin),data=data_case )
err<-residuals(mod)
N_nona<-N-sum(length(which(is.na(err))))
SSE=sum(err^2,na.rm=T)
ME=sum(abs(err),na.rm=T)/N_nona
MSE=SSE/N_nona+ME^2
table_res<-rbind(table_res, c(
  paste("BH sst"),
  as.character(formula(mod)),coeffs(mod),"", "",AIC(mod) , SSE, ME, MSE) )
modR<-list.append(modR, sst=mod)
lines(y,predict(mod),lwd=3,col=4)
legend_color<-c(legend_color,4)
legend_names<-c(legend_names, "BH sst")
legend_lty<-c(legend_lty,1)


# #so
  mod<-nls(R~a*S/(1+b*S)*exp(-c*so), start=list(a=ain,b=bin,c=cin),data=data_case )
  err<-residuals(mod)
  N_nona<-N-sum(length(which(is.na(err))))
  SSE=sum(err^2,na.rm=T)
  ME=sum(abs(err),na.rm=T)/N_nona
  MSE=SSE/N_nona+ME^2
  table_res<-rbind(table_res, c(
    paste("BH so"),
    as.character(formula(mod)),coeffs(mod),"", "",AIC(mod) , SSE, ME, MSE) )
  lines(y,predict(mod),lwd=3,col=5)
  legend_color<-c(legend_color,5)
  legend_names<-c(legend_names, "BH so")
  legend_lty<-c(legend_lty,1)
  
  
  # #botso
  mod<-nls(R~a*S/(1+b*S)*exp(-c*botso), start=list(a=ain,b=bin,c=cin),data=data_case )
  err<-residuals(mod)
  N_nona<-N-sum(length(which(is.na(err))))
  SSE=sum(err^2,na.rm=T)
  ME=sum(abs(err),na.rm=T)/N_nona
  MSE=SSE/N_nona+ME^2
  table_res<-rbind(table_res, c(
    paste("BH botso"),
    as.character(formula(mod)),coeffs(mod),"","", AIC(mod) , SSE, ME, MSE) )
  lines(y,predict(mod),lwd=3,col=6)
  legend_color<-c(legend_color,6)
  legend_names<-c(legend_names, "BH botso")
  legend_lty<-c(legend_lty,1)
  
   #nppv
  mod<-nls(R~ a*S/(1+b*S)*exp(-c*nppv), start=list(a=ain,b=bin,c=cin),data=data_case )
  err<-residuals(mod)
  N_nona<-N-sum(length(which(is.na(err))))
  SSE=sum(err^2,na.rm=T)
  ME=sum(abs(err),na.rm=T)/N_nona
  MSE=SSE/N_nona+ME^2
  table_res<-rbind(table_res, c(
    paste("BH nppv"),
    as.character(formula(mod)),coeffs(mod),"", "",AIC(mod) , SSE, ME, MSE) )
  lines(y,predict(mod),lwd=3,col=7)
  legend_color<-c(legend_color,7)
  legend_names<-c(legend_names, "BH nppv")
  legend_lty<-c(legend_lty,1)
  

legend("topright",legend=legend_names,col=legend_color,lwd=c(2,2),lty=legend_lty)

```

Now we will explore the models with two covariates.

```{r BH model with two covariates, echo=TRUE, warning = FALSE, message = FALSE}
plot(y,R,lwd=1,pch=19,xlab="Year",ylab="Recruits",ylim=c(min(R)*0.9,max(R)*2))
lines(y,predict(modR[[1]]),lwd=3,col="black")
title(paste(case, "-hindcast"))
legend_names<-"BH"
legend_color<-"black"
legend_lty<-1


#botT-nppv
mod<-nls(R~ a*S/(1+b*S)*exp(-c*botT-d*nppv), start=list(a=ain,b=bin,c=cin,d=din),data=data_case )
err<-residuals(mod)
N_nona<-N-sum(length(which(is.na(err))))
SSE=sum(err^2,na.rm=T)
ME=sum(abs(err),na.rm=T)/N_nona
MSE=SSE/N_nona+ME^2
table_res<-rbind(table_res, c(
  paste("BH botT-nppv"),
  as.character(formula(mod)),coeffs(mod),"",AIC(mod) , SSE, ME, MSE) )
modR<-list.append(modR, botTnppv=mod)
lines(y,predict(mod),lwd=3,col=5,lty=2)
legend_color<-c(legend_color,5)
legend_names<-c(legend_names, "BH botT-nppv")
legend_lty<-c(legend_lty,2)



#sst-nppv
mod<-nls(R~ a*S/(1+b*S)*exp(-c*sst-d*nppv), start=list(a=ain,b=bin,c=cin,d=din),data=data_case )
err<-residuals(mod)
N_nona<-N-sum(length(which(is.na(err))))
SSE=sum(err^2,na.rm=T)
ME=sum(abs(err),na.rm=T)/N_nona
MSE=SSE/N_nona+ME^2
table_res<-rbind(table_res, c(
  paste("BH sst-nppv"),
  as.character(formula(mod)),coeffs(mod),"", AIC(mod) , SSE, ME, MSE) )
modR<-list.append(modR, sstnppv=mod)
lines(y,predict(mod),lwd=3,col=8,lty=2)
legend_color<-c(legend_color,8)
legend_names<-c(legend_names, "BH sst-nppv")
legend_lty<-c(legend_lty,2)


# # #so-nppv 
  mod<-nls(R~a*S/(1+b*S)*exp(-c*so-d*nppv), start=list(a=ain,b=bin,c=cin,d=din),data=data_case )
  err<-residuals(mod)
  N_nona<-N-sum(length(which(is.na(err))))
  SSE=sum(err^2,na.rm=T)
  ME=sum(abs(err),na.rm=T)/N_nona
  MSE=SSE/N_nona+ME^2
  table_res<-rbind(table_res, c(
    paste("BH so-nppv"),
    as.character(formula(mod)),coeffs(mod),"", AIC(mod) , SSE, ME, MSE) )
  lines(y,predict(mod),lwd=3,col=9,lty=2)
  legend_color<-c(legend_color,9)
  legend_names<-c(legend_names, "BH so-nppv")
  legend_lty<-c(legend_lty,2)

  
  # #botso-nppv
  mod<-nls(R~ a*S/(1+b*S)*exp(-c*botso-d*nppv), start=list(a=ain,b=bin,c=cin,d=din),data=data_case )
  err<-residuals(mod)
  N_nona<-N-sum(length(which(is.na(err))))
  SSE=sum(err^2,na.rm=T)
  ME=sum(abs(err),na.rm=T)/N_nona
  MSE=SSE/N_nona+ME^2
  table_res<-rbind(table_res, c(
    paste("BH botso-nppv"),
    as.character(formula(mod)),coeffs(mod),"", AIC(mod) , SSE, ME, MSE) )
  lines(y,predict(mod),lwd=3,col=10,lty=2)
  legend_color<-c(legend_color,10)
  legend_names<-c(legend_names, "BH botso-nppv")
  legend_lty<-c(legend_lty,2)

legend("topright",legend=legend_names,col=legend_color,lwd=c(2,2),lty=legend_lty)

```

In this case we do not explore models with three covariates, because of the correlation among them.

# Summary table preparation

A draft summary table is here produced. Looking to the results, the model having the best predictive capability is the one depending only on SST.


```{r summary table, echo=TRUE, warning = FALSE, message = FALSE}

table_res<-as.data.frame(table_res)
colnames(table_res)<-c("Model","Formula","a","b","c","d","f","AIC","SSE","MAE","MSE")

table_res$a=as.numeric(table_res$a)
table_res$b=as.numeric(table_res$b)
table_res$c=as.numeric(table_res$c)
table_res$d=as.numeric(table_res$d)
table_res$f=as.numeric(table_res$f)
table_res$AIC=as.numeric(table_res$AIC)
table_res$SSE=as.numeric(table_res$SSE)
table_res$MAE=as.numeric(table_res$MAE)
table_res$MSE=as.numeric(table_res$MSE)

table_res$RMSE=sqrt(table_res$MSE)
table_res$rRMSE=table_res$RMSE/table_res[1,]$RMSE
table_res$rMAE=table_res$MAE/table_res[1,]$MAE

#table_res=rbind(names,table_res)
knitr::kable(table_res,digits=2)

```

# Conclusions

Note that here only the BevHolt model estimation was reported, but Ricker and hockey stick were also explored in SEAwise. 

Using the presented approach it is possible to explore any SRR model just adapting the script. It would be also possible to adapt the scipt to fit EMSRR applying a model error approach (in case a robust and contrasted data set is available).

More complex model performance indicators can be introduced following Marc presentation on model selection. 



